### 2018-02-28 thesis/contents/reviews/multiplesclerosis.tex
Updates from ShareLaTeX

\subsection{Pain and MS}
Pain is a common co-morbidity of MS, with up to 50-85\% prevalence \cite{Osterberg2005}. Common pain symptoms include acute pain such as optic neuritis and painful tonic seizure, as well as chronic pain such as trigeminal neuralgia. 91\% of the patients report pain at the time of evaluation, and most experience daily pain. Both central and peripheral pain are reports in people of MS. Central pain in MS is reported to occur 27-58\% of the subjects
\cite{OConnor2008}. The most common pain in MS include extremity pain and TN. TN has a prevalence 2\% in the MS, which is 20 times higher than in a general population.

TN secondary to MS, or symptomatic TN (MS-TN), is frequently indistinguishable from TN in terms of clinical presentation. It does however present with a demonstrable structural lesion other than vascular compression of CN V. MS-TN has a prevalence rate of about 2-4\% and in some studies, there are reports of prevalence rate as high as 6\%. Unilateral MS-TN is the most common pain type, however there is up to 30\% occurrences of bilateral pain.

>>>>

\subsection{Pain and MS}
Pain is a common co-morbidity of MS, with up to 50-85\% prevalence \cite{Osterberg2005}. Common pain symptoms include acute pain such as optic neuritis and painful tonic seizure, as well as chronic pain such as trigeminal neuralgia. 91\% of the patients report pain at the time of evaluation, and most experience daily pain. Both central and peripheral pain are reports in people of MS. Central pain in MS is reported to occur 27-58\% of the subjects \cite{OConnor2008}.

The most common pain in MS includes extremity pain and TN (MS-TN), which is also termed symptomatic TN. MS-TN has a prevalence 2\% in the MS, which is 20 times higher than in a general population \cite{Cruccu2009}. The clinical presentation of the pain is similar to classic TN. MS-TN has a prevalence rate of about 2-4\%, and in some studies, there are reports of prevalence rate as high as 6\%. Unilateral MS-TN is the most common pain type. However, there are up to 30\% occurrences of bilateral pain. MS-TN pathophysiology is believed to be different from classic TN, as it is often accompanied by the presentation of pontine MS lesions \cite{Cruccu2009}. The treatment of MS-TN via ablative interventions is also less effective than in classic TN \cite{Nurmikko2009}. The exact pathophysiological differences between MS-TN and classic TN, however, are not known. While it shares the feature of vascular compression with classic TN, its relevance to MS-TN pain has not been precisely determined \cite{Nurmikko2009}.
